# *Using Machine Learning to Predict Physical Exercise Practice Quality*
#### Project assignment of the Practical Machine Learning course from the
#### Johns Hopkins University's Data Science Specialization on Coursera.


### <u> *Background* </u>

Using devices such as Jawbone Up, Nike FuelBand, Fitbit, Garmin Vivofit, Microsoft Band, and Apple iWatch, among other products available in the market, it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the [PUC-Rio website](http://groupware.les.inf.puc-rio.br/har) (see the section on the Weight Lifting Exercise Dataset). 


### <u> *Loading and Cleaning the Testing and Training Data* </u>

The data for this project come from the [PUC Rio repository](http://groupware.les.inf.puc-rio.br/har). The two files in CSV format are: [training data](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv) and [test data](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv). I would like to thank the PUC Rio team for allowing their data to be used for this kind of assignment. 

```{r Loading.Cleaning.Data, cache=TRUE}

library(caret)
RawTestData <- read.csv("pml-testing.csv",na.strings=c("NA",""))
RawTrainData <- read.csv("pml-training.csv",na.strings=c("NA",""))
NAs <- apply(RawTrainData,2,function(x) {sum(is.na(x))}) 
CleanTrainData <- RawTrainData[,which(NAs == 0)]
CleanTestData <- RawTestData[,which(NAs == 0)]

```

In addition to the datasets above, a seed value (i.e. 1027) was used in this assignment to assure its reproducibility.

```{r seed.setting}

set.seed(1027)
options(warn=-1)

```

### <u> *Training and Cross Validation datasets* </u>

We used the code below to create the training and cross validation datasets.
We used 80% of the file for training and the remaining 20% for cross validation.

```{r Training.CrossValidation, cache=TRUE}

trainIndex <- createDataPartition(y = CleanTrainData$classe, p=0.8,list=FALSE)
trainSet <- CleanTrainData[trainIndex,]
crossValidationSet <- CleanTrainData[-trainIndex,]
# Unnecessary variables were removed using the code below
removeIndex <- as.integer(c(1,2,3,4,5,6))
trainSet <- trainSet[,-removeIndex]
testSet <- CleanTestData[,-removeIndex]

```


### <u> *Training code* </u>

The following training code was applied to generate the model.

```{r Training}

mytrControl = trainControl(method = "cv", number = 4)
modelFit <- train(trainSet$classe ~.,data = trainSet, method="rf", trControl = mytrControl)
modelFit

```


### <u> *Errors Calculation using the Cross Validation dataset* </u>

The errors were calculated using the cross validation dataset.

```{r ErrorCalc}

predicted <- predict(modelFit, crossValidationSet)
SampleError <- sum(predicted == crossValidationSet$classe)/nrow(crossValidationSet)

```


### <u> *Generating data for the prediction vector for the Assigment Submission* </u>

```{r prediction.vector.generation}

answers <- predict(modelFit, testSet)

```


### <u> *Generating the Answers and Answers files* </u>

Using the code provided by the PML course instructors we were able to generate 20 files with the answers there were uploaded in the second portion of the assignment.

```{r Answer.Files.Generation}

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(answers)

```

The answers are presented below:

```{r Answers}

answers

```


This model was able to predict all 20 answers correctly.
